---
title: "General Social Survey"
author: "Heather Laughlin"
date: "Year 2019"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
Gss <- read.spss("GSS2018.sav")
Gss1 <- as.tibble(Gss)
```

```{r include=FALSE,echo=FALSE}
#names(Gss1)
mydata <- select(Gss1,VOTE16,RACE,REGION,EDUC,BIGBANG,EVOLVED,DEGREE,POSTLIFE,AGE,OWNGUN,CAPPUN,COURTS,POLVIEWS,VOTE12,GOD)
mydata$AGE <- as.numeric(mydata$AGE)
mydata$EDUC <- as.numeric(mydata$EDUC)
mydata2 <-filter(mydata,RACE!="IAP" ) %>% droplevels()
mydata3 <- filter(mydata2,BIGBANG=="True" | BIGBANG=="False",EVOLVED=="True"|EVOLVED=="False",POSTLIFE=="YES"|POSTLIFE=="NO",OWNGUN=="YES"|OWNGUN=="NO",COURTS=="TOO HARSH"|COURTS=="NOT HARSH ENOUGH"|COURTS=="ABOUT RIGHT") %>% droplevels()
mydata4 <- filter(mydata3,CAPPUN=="FAVOR"|CAPPUN=="OPPOSE",VOTE12=="Voted"|VOTE12=="Did not vote",VOTE16=="Voted"|VOTE16=="Did not vote") %>% droplevels()
levels(mydata4$VOTE12)[1] <- "voted12"
levels(mydata4$VOTE12)[2] <- "did not vote in 12"

mydata5 <- filter(mydata4, CAPPUN=="FAVOR" | CAPPUN=="OPPOSE")

```



# Introduction Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This data is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 15 variables. It has further removed unwanted factor levels in much of the data. Examine that code to see what it is doing. First explain what you make out of comparing how people voted in 2012 with 2016.


How are the "voted in 2016"" and "voted in 2012"" related?

$H_0:$ voting in 2016 and 2012 are unrelated  
$H_A:$ voting in 2016 and 2012 are related 

We are comparing what happened with the people who voted in 2012 vs 2016. Do those who voted tend to always vote or does the election influence it? Did those who voted in 2012, still vote 2016 or is there a difference? Of those who did not vote in 2012, did they not vote in 2016? There is a noticable consistency of voters who didn't vote in 2012, not voting in 2016. But the voters have a larger shift from voting 2012 to not voting in 2016. Are these independent or are they related? 

## Methods


<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

We used the categorical variables, and a Chi-squared test to test the hypotheseis that voting in 2016 is / is not independent of voting in 2012

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

We used a barchart to show the voting trends between 2012 and 2016


#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~VOTE16+VOTE12, data=mydata4, type = "percent",xlab='voted in 2016') 

```


#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

Numerical summary is a cross tabulation:

```{r}
xtabs(~VOTE16+VOTE12, data=mydata4)

```



### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisqtestGC(~VOTE16+VOTE12, data=mydata4)
votetab=xtabs(~VOTE16+VOTE12, data=mydata4)
fisher.test(votetab)
```
$H_0:$ voting in 2016 is independent of voting in 2012  
$H_A:$ voting in 2016 is dependent of voting in 2012 

The hypothesis was that voting in 2012 is related to voting in 2016. If you voted in 2012 it is likely you voted in 2016. 

We did a chi-squared test to see if voting in 2016 was dependent on voting in 2012 results were 75.75. 

The p-value was< 2.2e-16, so we reject the null hypothesis which is smaller than .05.
Because the response and explanatory values are categorical, the bar plot and chi-squared test | test for independence.

# Introduction Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Now answer Is belief in the bigbang theory (BIGBANG) independent of belief in the afterlife (POSTLIFE)

We want to know if there is a correlation between belief in the "Bigbang" and belief in "Postlife".

We got our data from the 2018 General Social Survey.


## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

We used the categorical variables, and a Chi-squared test to test the hypotheseis that that belief in the "Bigbang"" is / is not independent of belief in "Postlife".

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

We used a barchart to show the data on belief in "Bigbang" and "Postlife".


#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->



```{r}
barchartGC(~BIGBANG+POSTLIFE, data=mydata3, type = "percent",xlab='BELIEVE IN POSTLIFE', main= 'Believe in BigBang')
```


#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}

xtabs(~BIGBANG+POSTLIFE, data=mydata3)

```


### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisqtestGC(~BIGBANG+POSTLIFE, data=mydata3)
votetab=xtabs(~BIGBANG+POSTLIFE, data=mydata3)
fisher.test(votetab)
```
$H_0:$ belief in "Bigbang" is independent on belief of "Postlife" 
$H_A:$ belief in "Bigbang" is dependent on belief of "Postlife"  

The null hypothesis was that belief in "Postlife" is unrelated to belief in "Bigbang".

There is a correlation between not believing in "Postlife" and believing on "Bigbang". there us a strong correlation between not believeing in "Postlife" and believing in "Bigbang". However, there is a larger percent of people who believe in both "Bigbang" and "Postlife". There are very few who do not believe in either "Bigbang" nor "Postlife"

If you believe in "Bigbang" it is likely you believe in "Postlife". We did a chi-squared test to see if "Postlife" was dependent on "Bigbang", the results was 7.4181. 

The p-value was 0.004074, so we reject the null hypothesis which is smaller than .05.

Because the response and explanatory values are categorical, the bar plot and chi-squared test | test for independence.




# Introduction Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This time compare the opinion on death penalty (CAPPUN) with opinion on courts (COURTS).

This data is from the 2018 General Social Survey. Is there a relationship between an opinion on the death penatly (CAPPUN) and the opinion on courts (COURTS).

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

We used the categorical variables, and a Chi-squared test to test the hypotheseis that that the opinion in the "CAPPUN"" is / is not independent of opinion in "COURTS".

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

We used a barchart to show the data on opinion on courts.

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~CAPPUN+COURTS, data=mydata5, type = "percent",xlab='VIEW ON CAPPUN', main= 'VIEW ON COURTS')
```


#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
xtabs(~CAPPUN+COURTS, data=mydata5)
```



### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisqtestGC(~CAPPUN+COURTS, data=mydata5)
votetab=xtabs(~CAPPUN+COURTS, data=mydata5)
fisher.test(votetab)
```
$H_0:$ There is no relationship between the view of the courts and the view on capital punishment. 
$H_A:$ There is a relationship between the view of the courts and the view on capital punishment. 

The p-value was  4.35e-10, so we reject the null hypothesis which is smaller than .05



# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

All three examples seemed to have some correlation in their variables. The 2012 voters mostly voted in 2016 and non's did not; it seemed that people tended to vote or not consistenetly. The non-belief in "POSTLIFE" seemed directly related to belief in the "BIGBANG". Though there was a larger group who believed in both the "BIGBANG" and "POSTLIFE". The "CAPPUN"went directly with the group who felt the "COURTS" were too lenient. 

